---
title: "Práctica 03"
output: github_document
---

## Load the following package: dplyr

```{r}
library("dplyr")
library("ggplot2")
```

## Load the data set carIns final. It already has the imputation of missing values.

```{r}
data <- load("../data/03_data/carIns_final.Rdata")
print(head(carIns_final,10))


```
## 1. Using the package dplyr, answer the following questions:
### (a) Obtain the number of cars by bodyStyle.

```{r}
# Obtener el número de autos por bodyStyle
resultado <- carIns_final  %>%  group_by(bodyStyle) %>% count()
print(resultado)
```
### (b) Obtain the number of cars by bodyStyle and fuelType.

```{r}
resultado <- carIns_final  %>%  group_by(bodyStyle, fuelType) %>% count()
print(resultado)

```
### (c) Obtain the mean and the standard deviation of the attribute cityMpg by bodyStyle in ascending order.
```{r}
resultado <- carIns_final %>%
  
# agrupa por bodyStyle  
  group_by(bodyStyle) %>%
# agrupa por bodyStyle  
#calculo de 
  summarize(mean_cityMpg = mean(cityMpg),
            sd_cityMpg = sd(cityMpg)) %>%
#orden ascendente
  arrange(mean_cityMpg)

print(resultado)
```

### (d) Also by bodyStyle, and for the attributes cityMpg and highwayMpg, obtain the mean, the standard deviation, the median and the inter-quartile range.
```{r}
resultado <- carIns_final %>% 
  group_by(bodyStyle) %>%
  summarize(
    mean_cityMpg = mean(cityMpg),
    sd_cityMpg = sd(cityMpg),
    median_cityMpg = median(cityMpg),
    IQR_cityMpg = IQR(cityMpg),
    
    mean_highwayMpg = mean(highwayMpg),
    sd_highwayMpg = sd(highwayMpg),
    median_highwayMpg = median(highwayMpg),
    IQR_highwayMpg = IQR(highwayMpg)
  )
resultado
```
# 2 Visualization
## 2. Using the package ggplot2, create graphs that you find adequate to answer the following questions.
### (e) Show the relationship between the attributes cityMpg and highwayMpg
```{r}
# Creacion de scatter plot
plot(carIns_final$cityMpg, carIns_final$highwayMpg, 
     xlab = "City MPG", ylab = "Highway MPG",
     main = "Relationship between City MPG and Highway MPG",
     pch = 16, col = "blue",
     xlim = c(0, 55), ylim = c(0, 55),
     xaxp = c(0, 50, 5), yaxp = c(0, 50, 5))

# Agregar linea de tendencia o trendline
abline(lm(carIns_final$highwayMpg ~ carIns_final$cityMpg), col = "red")

# Agregar lineas de cuadricula o gridlines
grid()

# Agregar una leyenda o legend
legend("top", legend = "Trendline", col = "red", lty = 1, bty = "n")

# Agregar etiquetas a algunos puntos de datos (opcional)
text(carIns_final$cityMpg[1:10], carIns_final$highwayMpg[1:10], 
     labels = rownames(carIns_final)[1:10], pos = 4, col = "black")

```

### (f) Show the distribution of cars by bodyStyle.
```{r}

# Count the number of cars by bodyStyle
bodyStyle_counts <- table(carIns_final$bodyStyle)

# Create a bar plot
barplot(bodyStyle_counts, 
        xlab = "Body Style",
        ylab = "Number of Cars",
        main = "Distribution of Cars by Body Style",
        col = "blue")

```

### (g) Show the distribution of cars by price. Suggestion: create bins of width equal to 5000.
```{r}

# Calcular los límites de los intervalos
min_price <- min(carIns_final$price)
max_price <- max(carIns_final$price)
interval_width <- 5000
breaks <- seq(min_price, max_price + interval_width, by = interval_width)

# Crear el gráfico de histograma
hist(carIns_final$price, 
     breaks = breaks,
     xlab = "Precio",
     ylab = "Número de Autos",
     main = "Distribución de Autos por Precio",
     col = "blue")


# USANDO GGPLOT2
# Crear los bins
interval_width <- 5000
breaks <- seq(min(carIns_final$price), max(carIns_final$price) + interval_width, by = interval_width)

# Crear el histograma usando ggplot2

ggplot(carIns_final, aes(x = price)) +
  
  geom_histogram(breaks = breaks, fill = "blue", color = "black") +
  
  labs(x = "Precio", y = "Numero de Carros",
       title = "Distribución de Autos por Precio") +
  theme_minimal()

```

### (h) Add the information of the density estimation to the previous graph.
```{r}
# Crear los bins
interval_width <- 5000
breaks <- seq(min(carIns_final$price), max(carIns_final$price) + interval_width, by = interval_width)

# Crear el histograma usando ggplot2

ggplot(carIns_final, aes(x = price)) +
  #geom_histogram(breaks = breaks, fill = "blue", color = "black", alpha = 0.5) +
  geom_density(color = "green", size = 3) +
    stat_density(geom = "line", color = "red", size = 1) +

  labs(x = "Precio", y = "Numero de Carros",
       title = "Distribución de Autos por Precio") +
  theme_minimal()


```

### (i) Check (visually) if it is plausible to consider that price follows a normal distribution.

```{r}

ggplot(carIns_final, aes(x = price)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(x = "Precio", y = "Densidad",
       title = "Distribución de Precios de Autos") +
  theme_minimal()

# Create Q-Q plot
# Crear el gráfico Q-Q utilizando qqnorm y qqline
qqnorm(carIns_final$price, main = "Gráfico Q-Q de Price", xlab = "Cuantiles teóricos", ylab = "Cuantiles observados")
qqline(carIns_final$price, col = "red")


# Calcular los límites de los intervalos
min_price <- min(carIns_final$price)
max_price <- max(carIns_final$price)
interval_width <- 5000
breaks <- seq(min_price, max_price + interval_width, by = interval_width)

# Crear los cuantiles teóricos y observados
quantiles_teoricos <- qnorm(ppoints(length(carIns_final$price)))
quantiles_observados <- quantile(carIns_final$price, probs = ppoints(length(carIns_final$price)))

# Crear el gráfico QQ utilizando ggplot2
library(ggplot2)

qq_data <- data.frame(QuantilesTeoricos = quantiles_teoricos, QuantilesObservados = quantiles_observados)

ggplot(qq_data, aes(x = QuantilesTeoricos, y = QuantilesObservados)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(x = "Cuantiles Teóricos", y = "Cuantiles Observados",
       title = "Gráfico QQ de carIns_final$price") +
  theme_minimal()

```

#### La distribución de los precios no sigue una distribucion normal

### (j) Show the distribution of price by make attribute. Suggestion: use boxplots and the function coord_flip().


```{r}
# Crear el gráfico de boxplots de la distribución de precios por atributo "make" con ggplot2


ggplot(carIns_final, aes(x = make, y = price)) +
  geom_boxplot(fill = "blue", color = "black", alpha = 0.5) +
  coord_flip() +
  labs(x = "Marca", y = "Precio",
       title = "Distribución de Precio por Marca") +
  theme_minimal()
```

### (k) Show the distribution of price by nDoors attribute. Suggestion: use histograms

```{r}

ggplot(carIns_final, aes(x = price, fill = nDoors)) +
  geom_histogram(binwidth = 5000, color = "black", alpha = 0.5) +
  labs(x = "Precio", y = "Número de Autos",
       title = "Distribución de Precio por Número de Puertas") +
  scale_fill_discrete(name = "Número de Puertas") +
  theme_minimal()

```

### (l) Show the distribution of price by bodyStyle and nDoors attributes. Suggestion: use histograms.

```{r}
ggplot(carIns_final, aes(x = price, fill = nDoors)) +
  geom_histogram(binwidth = 5000, position = "dodge", color = "black", alpha = 0.5) +
  facet_wrap(~ bodyStyle, nrow = 1) +
  labs(x = "Precio", y = "Número de Autos",
       title = "Distribución de Precio por Estilo de Carrocería y Número de Puertas") +
  scale_fill_discrete(name = "Número de puertas") +
  theme_minimal()




ggplot(carIns_final, aes(x = price)) +
  geom_histogram(binwidth = 5000, fill = "blue", color = "black", alpha = 0.5) +
  facet_grid(bodyStyle ~ nDoors) +
  labs(x = "Precio", y = "Número de Autos",
       title = "Distribución de Precio por Estilo de Carrocería y Número de Puertas") +
  theme_minimal()
```

### (m) Add the parameter scales="free_y" to the facet function in the previous graph.

```{r}
ggplot(carIns_final, aes(x = price, fill = nDoors)) +
  geom_histogram(binwidth = 5000, position = "dodge", color = "black", alpha = 0.5) +
  facet_wrap(~ bodyStyle, nrow = 1, scales = "free_y") +
  labs(x = "Precio", y = "Número de Autos",
       title = "Distribución de Precio por Estilo de Carrocería y Número de Puertas") +
  scale_fill_discrete(name = "Número de puertas") +
  theme_minimal()

ggplot(carIns_final, aes(x = price)) +
  geom_histogram(binwidth = 5000, fill = "blue", color = "black", alpha = 0.5) +
  facet_grid(bodyStyle ~ nDoors, scales = "free_y") +
  labs(x = "Precio", y = "Número de Autos",
       title = "Distribución de Precio por Estilo de Carrocería y Número de Puertas") +
  theme_minimal()

```

