---
title: "Práctica 02"
output: github_document
---
# 2 Hands On: Data Quality and Pre-Processing
## 1. Assessing Data Quality
### Load the following packages: dplyr, na.tools, tidyimpute (version from github decisionpatterns/tidyimpute”)
```{r}
library(na.tools)
library(devtools)
library(dplyr)
library(tidyverse)
library(tidyimpute)
library(sos)
library(ggplot2)
```

### Load the carInsurance data set about the insurance risk rating of cars based on several characteristics of each car
```{r}
#lee archivo csv, se indica que no contiene fila de encabezado
data <- read.csv('../data/02_data/carInsurance.data', header=FALSE)

#reemplaza valores ? por NA
data[data == "?"] <- NA

#crea un dataframe llamado carInsurance
#se usa el objeto data como argumento para la funcion data.frame()
carInsurance <- data.frame(data)

# vector de nombres de encabezado para el dataframe
header <- c("symboling","normalizedLosses","make","fuelType","aspiration","nDoors","bodyStyle","driveWheels","engineLocation","wheelBase","length","width","height","curbWeight","engineType","nCylinders","engineSize","fuelSystem","bore","stroke","compression-ratio","horsepower","peakRpm","cityMpg","highwayMpg","price")

# Establecer el encabezado al dataframe
names(carInsurance) <- header

#imprimir el dataframe, solo 10 filas
head(carInsurance,10)
```
### (a) Check if there are any missing values.


```{r}
columnas_valores_faltantes <- colSums(is.na(carInsurance))>0
#head(columnas_valores_faltantes,10)
nombres_columnas_valores_faltantes <- names(columnas_valores_faltantes[columnas_valores_faltantes])
nombres_columnas_valores_faltantes

valores_faltantes <- any_na(carInsurance)

#valores_faltantes

if (valores_faltantes) {
  print("Faltan valores en el conjunto de datos.")
} else {
  print("No hay valores faltantes en el conjunto de datos.")
}
```
### (b) Count the number of cases that have, at least, one missing value
```{r}
num_valores_faltantes <- carInsurance %>%
  filter_any_na() %>%
  count()
typeof(num_valores_faltantes)
print(num_valores_faltantes)

```
### (c) Create a new data set by removing all the cases that have missing values.
```{r}
#head(carInsurance,10)
#count(carInsurance)
carInsurance_sin_NA <- drop_rows_any_na(carInsurance)
head(carInsurance_sin_NA,10)
#count(carInsurance_sin_NA)
```
### (d) Create a new data set by imputing all the missing values with 0.
```{r}

# Crear una copia del conjunto de datos original y reemplaza los valores NA por ceros, se aplica a todo el conjunto de datos
carInsuranceImpute <- carInsurance %>% impute_zero_all()

# Reemplazar los valores NA por ceros en todas las columnas
#carInsuranceImpute <- impute_zero_all(carInsuranceImpute)

head(carInsuranceImpute, 10)
#count(carInsuranceImpute)


```
### (e) Create a new data set by imputing the mean in all the columns which have double type values.

```{r}



# Identificar columnas con valores de tipo double
columnas_double <- sapply(carInsuranceImpute, is.double)

#nombre de las columnas_double
nombres_columnas_double <- names(columnas_double[columnas_double])
#nombres_columnas_double
#typeof(nombres_columnas_double)

# Calcular la media de cada columna
media <- colMeans(carInsuranceImpute[,nombres_columnas_double])
media

# Crear un nuevo conjunto de datos imputando la media en las columnas correspondientes
carInsuranceMean <- carInsuranceImpute

for (col in nombres_columnas_double) {
  carInsuranceMean[[col]] <- media[col]
}
head(carInsuranceMean, 10)


```

### (f) Create a new data set by imputing the mode in all the columns which have integer type values.

```{r}
# Cargar el paquete necesario
library(modeest)


#identificar columnas int
columnas_integer <- sapply(carInsuranceMean, is.integer)

#print(columnas_integer)




#nombres de las columnas int
nombres_columnas_integer <- names(columnas_integer[columnas_integer])
#nombres_columnas_integer


# Crear una función personalizada para calcular la moda de un vector
moda_personalizada <- function(x) {
  moda <- unique(x)[which.max(tabulate(match(x, unique(x))))]
  if (length(moda) == 0) {  # Si no hay moda, devolver NA
    return(NA)
  } else {
    return(moda)
  }
}


# Calcular la moda en las columnas de tipo double
moda <- sapply(carInsuranceMean[, nombres_columnas_integer], moda_personalizada)

# Resultado: vector con las modas de cada columna
moda


# Crear un nuevo conjunto de datos imputando la media en las columnas correspondientes
carInsuranceMode <- carInsuranceMean 
head(carInsuranceMode, 10)


for (col in nombres_columnas_integer) {
  carInsuranceMode[[col]] <- moda[col]
}
head(carInsuranceMode, 10)

```


### (g) Create a new data set by imputing the most frequent value to the column ”nDoors”.
### Tip: use the function impute_replace()


```{r}



nDoorsMode <- names(which.max(table(carInsuranceMode$nDoors)))
nDoorsMode

carInsurancenDoorsMode <- carInsuranceMode %>%
  mutate(nDoors = impute_replace_all(nDoors, nDoorsMode))
head(carInsurancenDoorsMode, 10)

```

### (h) Combine the three last imputations to obtain a final dataset. Are there any duplicated cases?
### Tip: use the functions distinct() and count()


```{r}
# Verificar el resultado
newCarInsuranceFInal <- carInsurancenDoorsMode

# Verificar si existen casos duplicados
casosDuplicados <- newCarInsuranceFInal %>%
  distinct() %>%
  count()

# Imprimir el resultado
if (casosDuplicados$n == nrow(newCarInsuranceFInal)) {
  print("No hay casos duplicados en el conjunto de datos.")
} else {
  print("Existen casos duplicados en el conjunto de datos.")
}

# Verificar duplicados por columna
duplicadosPorColumna <- newCarInsuranceFInal %>%
  group_by(across(everything())) %>%
  count() %>%
  filter(n > 1) %>%
  select(-n)

# Imprimir el resultado
if (nrow(duplicadosPorColumna) == 0) {
  print("No hay columnas con duplicados en el conjunto de datos.")
} else {
  print("Columnas con duplicados y su conteo respectivo:")
  print(duplicadosPorColumna)
}

head(newCarInsuranceFInal,10)


```

## 2. Data Pre-Processing

### 2. Load the package dlookr. Use the same car insurance data set above and apply the following transformations to the price attribute. Be critical regarding the obtained results

```{r}
library(dlookr)


```

### (a) Apply range-based normalization and z-score normalization.
### Tip: use the function transform().


```{r}

newCarInsuranceFInal$price <- as.numeric(newCarInsuranceFInal$price)

newCarInsuranceFInal %>%
   mutate(rangeNorm = transform(newCarInsuranceFInal$price, method = "minmax"),
   zScoreNorm = transform(newCarInsuranceFInal$price, method = "zscore")) %>%
   select(rangeNorm, zScoreNorm) %>%
   head(20)

head(newCarInsuranceFInal,10)

```
### b) Discretize it into 4 equal-frequency ranges an into 4 equal-width ranges.
### Tip: use the function binning().
```{r}

library(dplyr)
library(binr)

newCarInsuranceFInal <- newCarInsuranceFInal %>%
  mutate(equalFrequency = binning(price, nbins = 4, type = "quantile"),
         equalWidth = binning(price, nbins = 4, type = "equal")) 
head(newCarInsuranceFInal,10)

# Imprimir los niveles de equalFrequency
cat("Niveles de equalFrequency: ", levels(newCarInsuranceFInal$equalFrequency), "\n")

# Imprimir los niveles de equalWidth
cat("Niveles de equalWidth: ", levels(newCarInsuranceFInal$equalWidth), "\n")

```

