---
title: "Práctica 03"
output: github_document
---

## Load the following package: dplyr

```{r}
library("dplyr")
library("ggplot2")
```

## Load the data set carIns final. It already has the imputation of missing values.

```{r}
data <- load("../data/03_data/carIns_final.Rdata")
#print(carIns_final)


```
## 1. Using the package dplyr, answer the following questions:
### (a) Obtain the number of cars by bodyStyle.

```{r}
# Obtener el número de autos por bodyStyle
resultado <- carIns_final  %>%  group_by(bodyStyle) %>% count()
print(resultado)
```
### (b) Obtain the number of cars by bodyStyle and fuelType.

```{r}
resultado <- carIns_final  %>%  group_by(bodyStyle, fuelType) %>% count()
print(resultado)

```
### (c) Obtain the mean and the standard deviation of the attribute cityMpg by bodyStyle in ascending order.
```{r}
resultado <- carIns_final %>%
  
# agrupa por bodyStyle  
  group_by(bodyStyle) %>%
# agrupa por bodyStyle  
#calculo de 
  summarize(mean_cityMpg = mean(cityMpg),
            sd_cityMpg = sd(cityMpg)) %>%
#orden ascendente
  arrange(mean_cityMpg)

print(resultado)
```

### (d) Also by bodyStyle, and for the attributes cityMpg and highwayMpg, obtain the mean, the standard deviation, the median and the inter-quartile range.
```{r}
resultado <- carIns_final %>% 
  group_by(bodyStyle) %>%
  summarize(
    mean_cityMpg = mean(cityMpg),
    sd_cityMpg = sd(cityMpg),
    median_cityMpg = median(cityMpg),
    IQR_cityMpg = IQR(cityMpg),
    
    mean_highwayMpg = mean(highwayMpg),
    sd_highwayMpg = sd(highwayMpg),
    median_highwayMpg = median(highwayMpg),
    IQR_highwayMpg = IQR(highwayMpg)
  )
resultado
```
# 2 Visualization
## 2. Using the package ggplot2, create graphs that you find adequate to answer the following questions.
### (e) Show the relationship between the attributes cityMpg and highwayMpg
```{r}
# Creacion de scatter plot
plot(carIns_final$cityMpg, carIns_final$highwayMpg, 
     xlab = "City MPG", ylab = "Highway MPG",
     main = "Relationship between City MPG and Highway MPG",
     pch = 16, col = "blue",
     xlim = c(0, 55), ylim = c(0, 55),
     xaxp = c(0, 50, 5), yaxp = c(0, 50, 5))

# Agregar linea de tendencia o trendline
abline(lm(carIns_final$highwayMpg ~ carIns_final$cityMpg), col = "red")

# Agregar lineas de cuadricula o gridlines
grid()

# Agregar una leyenda o legend
legend("top", legend = "Trendline", col = "red", lty = 1, bty = "n")

# Agregar etiquetas a algunos puntos de datos (opcional)
text(carIns_final$cityMpg[1:10], carIns_final$highwayMpg[1:10], 
     labels = rownames(carIns_final)[1:10], pos = 4, col = "black")

```
### (f) Show the distribution of cars by bodyStyle.
```{r}

# Count the number of cars by bodyStyle
bodyStyle_counts <- table(carIns_final$bodyStyle)

# Create a bar plot
barplot(bodyStyle_counts, 
        xlab = "Body Style",
        ylab = "Number of Cars",
        main = "Distribution of Cars by Body Style",
        col = "blue")

```

### (g) Show the distribution of cars by price. Suggestion: create bins of width equal to 5000.
```{r}

# Calcular los límites de los intervalos
min_price <- min(carIns_final$price)
max_price <- max(carIns_final$price)
interval_width <- 5000
breaks <- seq(min_price, max_price + interval_width, by = interval_width)

# Crear el gráfico de histograma
hist(carIns_final$price, 
     breaks = breaks,
     xlab = "Precio",
     ylab = "Número de Autos",
     main = "Distribución de Autos por Precio",
     col = "blue")


```
### (h) Add the information of the density estimation to the previous graph.
```{r}

```

