---
title: "Práctica 02"
output: github_document
---
## 1. Assessing Data Quality
## Load the following packages: dplyr, na.tools, tidyimpute (version from github decisionpatterns/tidyimpute”)
```{r}
library(dplyr)
library(na.tools)
library(tidyimpute)
```

## Load the carInsurance data set about the insurance risk rating of cars based on several characteristics of each car
```{r}

data <- read.csv('../data/01_datamanip/carInsurance.data', header=FALSE)

data[data == "?"] <- NA

carInsurance <- data.frame(data)

# List of names obtained of webpage
header <- c("symboling","normalizedLosses","make","fuelType","aspiration","nDoors","bodyStyle","driveWheels","engineLocation","wheelBase","length","width","height","curbWeight","engineType","nCylinders","engineSize","fuelSystem","bore","stroke","compression-ratio","horsepower","peakRpm","cityMpg","highwayMpg","price")

# Assign header a names of df
names(carInsurance) <- header

head(carInsurance,10)
```
### (a) Check if there are any missing values.


```{r}
has_missing <- any_na(carInsurance)

if (has_missing) {
  print("There are missing values in the dataset.")
} else {
  print("There are no missing values in the dataset.")
}
```
### (b) Count the number of cases that have, at least, one missing value
```{r}
# Count the number of cases with at least one missing value
num_missing_cases <- carInsurance %>%
  filter_any_na() %>%
  count()

print(num_missing_cases)

```
## (c) Create a new data set by removing all the cases that have missing values.
```{r}
carInsurance_sin_faltantes <- drop_rows_any_na(carInsurance)
head(carInsurance_sin_faltantes,10)

```
## (d) Create a new data set by imputing all the missing values with 0.
```{r}


library(tidyimpute)

# Crear una copia del conjunto de datos original
newCarInsurance <- carInsurance

# Reemplazar los valores NA por ceros en todas las columnas
newCarInsurance <- impute_zero_all(newCarInsurance)

head(newCarInsurance, 10)



```
## 
